---
title: "Getting started with ggpatina"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ggpatina}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ggpatina)
library(ggplot2)
```

# Introduction

`ggpatina` is a small toolkit for giving your **ggplot2** figures an analog
makeover. Effects are applied *after* the plot is rendered: you build a normal
plot, convert it to a [`magick`](https://docs.ropensci.org/magick/) image with
`as_magick()`, and then pass that image through one or more patina filters.
The package also includes a few lightweight themes and helpers for period
fonts so that the whole graphic carries the vibe. Code chunks below are
evaluated when the vignette is built, so any plots they produce will show up
automatically. This vignette walks through several effects to illustrate the
typical workflow.

# Blueprint style

We'll start with a basic scatter plot and gradually turn it into a blueprint.

```{r blueprint}
# build the plot as usual
base <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point()

# swap in the blueprint theme
p_bp <- base + theme_blueprint()

# convert to a magick image and apply the blueprint patina
img <- as_magick(p_bp)
patina_blueprint(img)
```

# Slide transparency

The transparency slide effect mimics a projected 35mm slide. We build on the
same base plot so you can see the incremental steps.

```{r slide}
# start from the base scatter plot
base <- ggplot(mtcars, aes(wt, mpg)) + geom_point(color = "steelblue")

# add the slide-friendly theme
p_slide <- base + theme_transparency()

# render to a magick image, then give it slide artifacts
img <- as_magick(p_slide)
slideify_transparency(img)
```

# Ink bleed

Ink bleed gives a photocopied or cheaply printed feel where edges blur and ink
spreads into the paper.

```{r bleed}
# reuse the base scatter plot
base <- ggplot(mtcars, aes(wt, mpg)) + geom_point()

# newspaper theme sets up the flat tones
p_news <- base + theme_newspaper()

# convert and let the ink bleed a bit
img <- as_magick(p_news)
patina_ink_bleed(img)
```

### Kodachrome photo

Warm colors and a soft fade evoke the look of vintage Kodachrome film.

```{r kodachrome}
base <- ggplot(mtcars, aes(wt, mpg)) + geom_point(color = "firebrick")
p_koda <- base + theme_transparency()
img <- as_magick(p_koda)
patina_kodachrome(img)
```

### Newscast screenshot

This effect mimics a low-resolution television capture.

```{r newscast}
base <- ggplot(mtcars, aes(wt, mpg)) + geom_point(color = "goldenrod")
p_cast <- base + theme_newscast()
img <- as_magick(p_cast)
patina_newscast(img)
```

### Photocopy

Misaligned colors and grain suggest a worn-out photocopier.

```{r photocopy}
base <- ggplot(mtcars, aes(wt, mpg)) + geom_point()
img <- as_magick(base)
patina_photocopy(img)
```

These examples are just a starting point. Combine effects and themes to give
your plots any era's personality.

